---
title: "Final project R"
author: "Ahmed Tarek Sayed"
date: "February 12, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

##importing the datasets
```{r}
tmdb_movies_metadata <- read_csv("E:/Academic/ITI/R/Evaluation Project/R for Data Science course evaluation/tmdb_movies_metadata.csv")

tmdb_cast_crew <- read_csv("E:/Academic/ITI/R/Evaluation Project/R for Data Science course evaluation/tmdb_cast_crew.csv")
```

### Is there a relationship between movie the number of budget and the movie rating?

```{r}
tmdb_movies_metadata %>% select(budget, vote_average)%>% summary()
```
the mean is far away from median indicating the variable is right skewed, indicating the presence of extremes or ouytliers in the budget variable
however the voting is balanced.

```{r}
tmdb_movies_metadata %>% ggplot() + geom_freqpoly(mapping = aes(x=budget ))
```
there is clearly very large number of movies with budget = 0!!
Iwill devide the budget into 20 equal parts to try to see if the 0 budget movies are an entry (null) or not.
```{r}
tmdb_movies_metadata %>% mutate(percentile = ntile(budget,20)) %>% select(percentile, budget,vote_average,vote_count) %>% group_by(percentile) %>% summarise(n(), mean(budget),min(budget),median(budget), max(budget), mean(vote_average), mean(vote_count) )
```

I can see normal equal average voting rate in all 20 parts regardless of the 0 budget, so I will deal with them as an entry error.

all 20 parts are almost normally dist with mean = media except in the 5th which is the break point between the 0 budget(entery error) and the non 0, so i will try to get the best number that separate the real data from the entery error 
```{r}
tmdb_movies_metadata %>%filter(budget < 100000) %>% mutate(percentile = ntile(budget,30)) %>% select(percentile, budget,vote_average,vote_count) %>% group_by(percentile) %>% summarise(n(), mean(budget),min(budget),median(budget), max(budget) )
```

Now i will see the movies with budget between 1 and 22000 to see if they make sense or not
```{r}
tmdb_movies_metadata %>%filter(budget %in% (1:50000)) %>% select(title, release_date, budget, vote_average, vote_count, popularity, runtime) %>%  arrange(budget)
```

Can old movies budget be that small?
I have checked the IMDP data base to see if even the old movies <1940 budget can make sense with very low numbers but my search revealed error with these entery 
examples of errors compared to imdb:-
http://www.imdb.com/title/tt0022879/ dataset:4 imdb:800K
http://www.imdb.com/title/tt1714210/ dataset:8000 imdb:120K

Idecided to take >8000 as the limit of real non outliers data
```{r}
filtered_movies <- tmdb_movies_metadata %>% filter(budget >8000)
```

Trying to visualise the relation bettween voting average and budget.

```{r}
filtered_movies %>% ggplot(aes(x=budget, y=vote_average)) + geom_point() + geom_smooth()
```
It seems that there is no correlation.
My interpertation : we have 2 problems :-
1- we cannot compare budget from diffrent years to each other
2- voting rate can be really missleading specially with with low voting count

I will extract the year of release from date
```{r}
filtered_movies <- filtered_movies %>% mutate(release_year = format(release_date,"%Y"))

```

```{r}
filtered_movies %>% ggplot(aes(x=vote_average, y=budget)) + geom_point(aes(color = release_year),show.legend = F) + geom_smooth()
```
I have downloaded cpi dataset from  https://datahub.io/core/cpi-us#data and inner joined it to movie dataset to be able to adjust budget per year
```{r}
cpiai_csv <- read_csv("E:/Academic/ITI/R/Evaluation Project/R for Data Science course evaluation/cpiai_csv.csv", 
    col_types = cols(Date = col_date(format = "%d/%m/%Y")))

cpi <- cpiai_csv %>% transmute(release_year = format(Date,"%Y"), cpi= Index) 

cpi <- cpi %>%  group_by(release_year) %>% summarise( cpi = mean(cpi, na.rm = T))

indexed_movies <- cpi %>% select(release_year, cpi) %>% right_join(filtered_movies, by = "release_year")
```
I have got the mean cpi for the year and then right joind the cpi to the movie dataset with the year 
```{r}
indexed_movies <- indexed_movies %>% mutate( indexed_budget = budget/cpi)
```
I have created a new variable called the indexed_budget revealing the budget in relation to yearly index

Now lets examine the relation between the total_votes(count*average) againist the indexed budget

```{r}
indexed_movies %>% mutate(total_votes = vote_count*vote_average) %>%  ggplot(aes(x=total_votes, y=indexed_budget)) + geom_point(mapping = aes(color = release_year), show.legend = F) + geom_smooth() + ylim(NA,2000000)
```
there is a moderate correlation between the two but from the colors we can see that there is a clear groupping by the year.
the voting count is increasing throw the years making this correlation noised by the effect of count per time

You may notice a very clear outlier with indexed budget >4000000 (very extreme).

```{r}
indexed_movies %>% filter(indexed_budget > 4000000)
```
after checking the movie at imdb 
http://www.imdb.com/title/tt0017136/ dataset:92m imdb:6m (clearly an error)

###What are the genres that have the highest average rating?

First of all i need to know the max number of genres per movie 
```{r}
indexed_movies %>% transmute(count_gen = str_count(genres, fixed("|")) +1) %>% summarise(max(count_gen, na.rm = T))
```
Then I am going to separate these words
```{r}
spr_gen_movies <- indexed_movies %>% separate(genres, into = paste("gen", 1:7, sep = "_"))
```
Now I amgoing to gather them and filter all the nulls to make a group function

```{r}
gen_movies <- spr_gen_movies %>%  gather(num_range("gen_", 1:7), key = "gen_no", value = "gen") %>% filter(!is.na(gen))
```

Now lets play :D 
```{r}
gen_movies <- gen_movies %>% mutate(total_votes = vote_count*vote_average)
gen_movies %>% group_by(gen) %>% summarise(avg_rate = mean(vote_average), avg_total_votes = mean(total_votes)) %>% arrange(desc(avg_rate)) 
```
it looks like war then history then forign movies is the one with highest avg_rate
but why do I feel like it is not the truth, You dont see alot of fans of war or forign movies :D 
but lets see the total_votes(count*average) as it is a more accurate indicator.

```{r}
gen_movies <- gen_movies %>% mutate(total_votes = vote_count*vote_average)
gen_movies %>% group_by(gen) %>% summarise(avg_rate = mean(vote_average), avg_total_votes = mean(total_votes)) %>% arrange(desc(avg_total_votes)) 
```
Now the results are more logical, the top 4 genres are fiction, scince, adventure then fantasy
looks like the next part of the avenger series going to be a hit :D

###What are the plot keywords that have the highest average rating?

first I need to count the maximum number of keywords in single row 
```{r}
indexed_movies %>% transmute(count_key = str_count(keywords, fixed("|")) +1) %>% summarise(max(count_key, na.rm = T))
```

Now I am going separate the column into 97 column(max number of keywords per row)
```{r}
spr_key_movies <- indexed_movies %>% separate(keywords, into = paste("key", 1:97, sep = "_"))
```
 Then I am going to gather al columns in on column by repeating records and removing the nulls to be able to make group function 
```{r}
key_movies <- spr_key_movies %>%  gather(num_range("key_", 1:97), key = "key_no", value = "key") %>% filter(!is.na(key))
```

I have made groupping by the keywords and summarised the mean of the avg rating and also my total_vote(calculated) 
```{r}
key_movies <- key_movies %>% mutate(total_votes = vote_count*vote_average)

key_movies %>% group_by(key) %>% summarise(avg_rate = mean(vote_average), avg_total_votes = mean(total_votes)) %>% arrange(desc(avg_rate)) 

key_movies %>% group_by(key) %>% summarise(avg_rate = mean(vote_average), avg_total_votes = mean(total_votes)) %>% arrange(desc(avg_total_votes))
```
Using the vote_avg shows that  conservatory, heirloom then kamikaze are the key words with the highst rate

##Who are the highly rated directors? Who are the highly rated actors?
In order to see the highly rated director I need to join the 2 tables of metadata and crew
and why dont we also calculate the (vote_count * vote_average) to see which KPI is more accurate

```{r}
full_movies <- left_join(indexed_movies, tmdb_cast_crew, by = "movie_id")
full_movies <- full_movies %>% mutate(total_votes = vote_count*vote_average)
```
now lets see the director with the highest rating averrage

```{r}
full_movies %>% group_by(director) %>% summarise(sum_vote = sum(vote_average), sum_total = sum(total_votes)) %>%  arrange(desc(sum_vote))

full_movies %>% group_by(director) %>% summarise(sum_vote = sum(vote_average), sum_total = sum(total_votes)) %>%  arrange(desc(sum_total))
```
I have arranged the list both by sum of vote average (using the mean give very poor accuracy) and sum total_votes.
The total votes KPI shows to be more accurate than vote averrage when compared to imdb(one can know which is better just by seeing Christopher Nolan at the top :D)
Refrence : http://www.imdb.com/list/ls069174126/

1regarding the best actor:-
I have gathered all the acttors in one value column 'actor' and removed any null columns 
```{r}
actor_movie <- full_movies %>%
    gather(key = actor_no, value = actor, starts_with('actor'), na.rm = T)
```

now lets see using our KPIs
```{r}
actor_movie %>% group_by(actor) %>% summarise(sum_vote = sum(vote_average), sum_total = sum(total_votes)) %>%  arrange(desc(sum_vote))

actor_movie %>% group_by(actor) %>% summarise(mean_vote = mean(vote_average), sum_total = sum(total_votes)) %>%  arrange(desc(sum_total))
```

well now both give moderate reults now(I think it is up to ones taste now :D)

##Who are the most profitable directors? Who are the most profitable actors?
In order to know profit we need to calc it (profit = revenue - cost(budget)) then I am going to devide by the (consumer price index) to be able to compare diffrent times

```{r}
full_movies %>% mutate(profit = revenue - budget, indexed_profit = profit/cpi) %>% group_by(director) %>% summarise(sum_profit = sum(profit), sum_indexed_profit = sum(indexed_profit)) %>% arrange(desc(sum_profit)) 

full_movies %>% mutate(profit = revenue - budget, indexed_profit = profit/cpi) %>% group_by(director) %>% summarise(sum_profit = sum(profit), sum_indexed_profit = sum(indexed_profit)) %>% arrange(desc(sum_indexed_profit)) 
```

Well without putting out the inflation the most profitable directors are steven spilberg, peter jackson, james cameron

but after putting out inflation() the list changes a little as directors who made old profitable movies are given a higher weight

Regarding actors
```{r}
actor_movie %>% mutate(profit = revenue - budget, indexed_profit = profit/cpi) %>% group_by(actor) %>% summarise(sum_profit = sum(profit), sum_indexed_profit = sum(indexed_profit, na.rm = T)) %>% arrange(desc(sum_profit)) 

actor_movie %>% mutate(profit = revenue - budget, indexed_profit = profit/cpi) %>% group_by(actor) %>% summarise(sum_profit = sum(profit), sum_indexed_profit = sum(indexed_profit,na.rm = T)) %>% arrange(desc(sum_indexed_profit)) 
```
also the first list shows Robert downey jr , tom cruise, tom hanks

but using the cpi we can reveal old actors who were a hit in there times  such as Vivien Leigh


